### 2024-10-02 paper/proposal.qmd
set up initial files

---
title: "STA302 Final Project Proposal"
subtitle: ""
author: 
  - name: "Ziqi Zhu"
  - name: "Author 2"
  - name: "Author 3"
thanks: "Code and data are available at: https://github.com/zzq20010617/sta302Paper"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
date-format: long
abstract: ""
number-sections: true
bibliography: proposal_references.bib
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
---

```{r}
#| include: false
#| warning: false
#| message: false

library(ggplot2)

```

# Introduction (350)

# Data description (300)
The R programming language (@talia) and ggplot2（@ggplot2）were used for processing the data.

# Ethics discussion (100-200)

# Preliminary results (300)

\newpage

# References