### 2024-10-11 paper/proposal.qmd
add introduction

---
title: "STA302 Final Project Proposal"
subtitle: ""
author: 
  - name: "Ziqi Zhu"
  - name: "Author 2"
  - name: "Author 3"
thanks: "Code and data are available at: https://github.com/zzq20010617/sta302Paper"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
date-format: long
abstract: ""
number-sections: true
bibliography: proposal_references.bib
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
---

```{r}
#| include: false
#| warning: false
#| message: false

library(ggplot2)
library(readr)
library(tibble)
library(knitr)
library(kableExtra)
library(dplyr)
data <- read.csv("../data/analysis_data/analysis_data.csv")
```

# Introduction (350)
The gaming industry has experienced exponential growth in recent years, with millions of players engaging in various games on platforms like Steam. Understanding what influences average playtime in games is crucial for game developers, publishers, and marketers. Average playtime is a key indicator of player engagement and game success, and investigating factors that impact it can provide valuable insights into game design, pricing strategies, and marketing efforts. So our research aims to explore the factors that influence the average playtime of games available on Steam using a dataset we get from Kaggle. Specifically, we will investigate how variables such as price, user ratings, and other game characteristics contribute to player engagement, as measured by playtime.
From some previous papers, we learned that both positive and negative reviews have a significant correlation with playtime, but the effect is more pronounced for positive reviews(@paper1). Another paper found out that game pricing significantly affects player engagement, with lower-priced games generally seeing higher playtime.(@paper2). The third paper did an empirical study that also examined the influence of game reviews on playtime, focusing on how the number and sentiment of reviews can affect player engagement. It concludes that games with more positive reviews see increased playtime, especially if those reviews highlight the game’s quality(@paper3).
We found this problem suitable for linear regression because it allows us to model the relationship between average playtime (the dependent variable) and various predictors (price, ratings, developer, etc.). This approach helps quantify how each factor influences playtime, enabling a clear understanding of the impact of different game attributes. By fitting a linear regression model, we can determine which factors have statistically significant effects on playtime, providing actionable insights for game developers and marketers.

# Data description (300)
We get the Steam games data(@davis2023steam) from Kaggle. The R programming language (@talia) and packages readr(@readr), httr(@httr), jsonlite(@jsonlite) were used to download the data, dplyr(@dplyr) and lubridate(@lubridate) were used to clean the data. Data was originally gathered from the Steam Store and SteamSpy APIs around May 2019. This table @tbl-cle shows the first 3 entries of cleaned data. We are going to take average_playtime as the response variable, it measures the mean time (in minutes) that players spend on a game, a summary of this variable is shown in @tbl-aveplaytime. This variable captures overall player engagement and serves as a good indicator of how immerse or entertaining a game is. It is suitable for a linear regression model because it is continuous and quantitative. The predictors selected are price, positive ratings, negative ratings, developer, and number of owners. All are numeric except for the number of owners, which is categorical since it represents estimated ranges. Price may influence playtime, as higher costs could lead to longer engagement. Positive ratings and negative ratings are reviews of players of that game and it can only be positive or negative. Positive ratings likely correlate with greater playtime, while negative ratings could indicate the opposite. Developer refers to the development company of the game, which could affect playtime, with established studios often producing longer, high-quality games. Lastly, the number of owners is an estimated number of owners, containing lower and upper bounds (like 20000-50000), it could signal popularity, where more owners may suggest higher average playtime. Summary of numerical predictors can be find in table @tbl-predictor.
```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-cle
#| tbl-cap: "Steam Games Data"
#| tbl-float: false

head(data, 3) %>% 
  kbl("latex") %>%
  kable_styling(latex_options = c("striped", "scale_down")) 
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-aveplaytime
#| fig-pos: "H"

summary_table <- data %>%
  summarise(
    Mean = round(mean(average_playtime, na.rm = TRUE), 2),
    Median = round(median(average_playtime, na.rm = TRUE), 2),
    Std_Dev = round(sd(average_playtime, na.rm = TRUE), 2),
    Min = round(min(average_playtime, na.rm = TRUE), 2),
    Max = round(max(average_playtime, na.rm = TRUE), 2),
  )
kable(summary_table, caption = "Summary of Average Playtime")
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-predictor

predict_summary_table <- data %>%
  summarise(
    mean_price = round(mean(price, na.rm = TRUE), 2),
    median_price = round(median(price, na.rm = TRUE), 2),
    mean_positive_ratings = round(mean(positive_ratings, na.rm = TRUE), 2),
    median_positive_ratings = round(median(positive_ratings, na.rm = TRUE), 2),
    mean_negative_ratings = round(mean(negative_ratings, na.rm = TRUE), 2),
    median_negative_ratings = round(median(negative_ratings, na.rm = TRUE), 2),
  )
kable(predict_summary_table, caption = "Summary of Predictors") %>%
kable_styling(latex_options = c("striped", "scale_down"), font_size = 6) 
```

# Ethics discussion (100-200)

# Preliminary results (300)

\newpage

# References