### 2024-12-01 paper/paper.Rmd
update paper, save models in rds files

---
title: "Exploring Factors Influencing Average Playtime of Video Games: A Regression Analysis"
subtitle: ""
author: 
  - name: "Ziqi Zhu"
  - name: "Raghav Sinha"
  - name: "Christophe McWatt"
thanks: "Code and data are available at: [https://github.com/zzq20010617/sta302Paper](https://github.com/zzq20010617/sta302Paper)."
date: today
date-format: long
abstract: ""
number-sections: true
bibliography: references.bib
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
---

```{r}
#| include: false
#| warning: false
#| message: false

library(ggplot2)
library(GGally)
library(readr)
library(tibble)
library(knitr)
library(kableExtra)
library(dplyr)
library(MASS)
library(caret)
library(car)
library(broom)
```

```{r}
#| include: false
#| warning: false
#| message: false

# Load analysis data
data <- read.csv(here::here("data/analysis_data/analysis_data.csv"))
```

# Contributions
We found the dataset together, and seperate the works as follow:
Ziqi: Introduction and data description
Raghav: Ethics discussion and some plot in preliminary result
Christophe: Most of Preliminary results section

# Introduction
The gaming industry has experienced exponential growth in recent years, with millions of players engaging in various games on platforms like Steam. Understanding what influences average playtime in games is crucial for game developers, publishers, and marketers. Average playtime is a key indicator of player engagement and game success, and investigating factors that impact it can provide valuable insights into game design, pricing strategies, and marketing efforts. So our research aims to explore the factors that influence the average playtime of games available on Steam using a dataset we get from Kaggle. Specifically, we will investigate how variables such as price, user ratings, and other game characteristics contribute to player engagement, as measured by playtime.
From some previous papers, we learned that both positive and negative reviews have a significant correlation with playtime, but the effect is more pronounced for positive reviews [@paper1]. Another paper found out that game pricing significantly affects player engagement, with lower-priced games generally seeing higher playtime [@paper2]. The third paper did an empirical study that also examined the influence of game reviews on playtime, focusing on how the number and sentiment of reviews can affect player engagement. It concludes that games with more positive reviews see increased playtime, especially if those reviews highlight the game’s quality [@paper3].
We found this problem suitable for linear regression because it allows us to model the relationship between average playtime (the dependent variable) and various predictors (price, ratings, developer, etc.). This approach helps quantify how each factor influences playtime, enabling a clear understanding of the impact of different game attributes. By fitting a linear regression model, we can determine which factors have statistically significant effects on playtime, providing actionable insights for game developers and marketers.

# Methods

To develop and refine the final model for predicting average playtime, a structured approach was employed that combined diagnostic assessments, transformations, and model selection techniques. The analysis began with data preparation and selection of relevant predictors based on domain knowledge and preliminary exploration. A multiple linear regression model was initially fitted using predictors including price, positive ratings, negative ratings, number of achievements as numeric variables, and ownership levels as categorical variable.

Preliminary model diagnostics, since we are using MLR model, we first need to check conditional mean condition with both response and predictors, including response vs. fitted plots and pairwise plots between predictors. Then residuals vs. fitted plots were used to assess the assumptions of linearity, constant-variance, and uncorrelated errors. While Normal Quantile-Quantile (QQ) plot is used to check for normality. To address violations of these assumptions, Box-Cox transformations were applied to the response variable and select predictors to make the most suited transform function based on their $\lambda$. Then pairwise scatterplots and the Variance Inflation Factor (VIF) were used to check for multicollinearity, and highly correlated variables were removed or adjusted to improve the model's stability.

Refinement of the model was carried out through stepwise selection using AIC (Akaike Information Criterion) to identify the best combination of predictors while balancing model complexity and fit. ANOVA tests were used to compare nested models and evaluate whether reductions in predictors led to significant changes in explanatory power. The final model was validated using diagnostic plots to confirm adherence to linear regression assumptions, ensuring that it was suitable for interpretation and prediction.

The methodology was implemented in programming language **R** [@citeR]. Packages **readr** [@readr], **httr** [@httr], **jsonlite** [@jsonlite] were used to download the data, **dplyr** [@dplyr] and **lubridate** [@lubridate] were used to clean the data. The diagnostic process is done with help of **ggplot2** [@ggplot2] for residual graphs, **GGally** [@GGally] and **car** [@car] for multicollinearity checks, **MASS** [@MASS] for stepwise selection, **caret** [@caret] is used to perform box-cox transform on variables. This systematic approach ensured that each decision was guided by statistical evidence and diagnostic tools, allowing for the development of a robust final model. A simplified version of the methodology flowchart is presented in @fig-flowchart.

```{r}
#| warning: false
#| message: false
#| echo: false
#| out.width: 100%
#| label: fig-flowchart
#| fig-cap: "Methodology flowchart"

include_graphics(here::here("other/flowchart.png"))
```

<!-- ## Fit multiple linear regression model -->
```{r}
#| echo: false
#| warning: false
#| message: false

# Load first model fit with lm(average_playtime ~ price + positive_ratings + 
# negative_ratings + achievements + owners, data=analysis_data)
model_1 <- readRDS(file = here::here("models/model_1.rds"))
```
# Results

At the starting point, our baseline model uses `average_playtime` as the response variable, with `price`, `positive_ratings`, `negative_ratings`, `achievements`, and `owners` as predictors. The four key linear regression assumptions considered are linearity, uncorrelated errors, constant error variance, and normality. Additionally, since this is a multiple linear regression model, we also examine the conditional mean response and predictor assumptions.
<!-- ## Create Response vs Fitted Scatterplot to check patterns in residuals before further analysis. -->
```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-resp-fit
#| fig-cap: "Response vs. Fitted plot to assess model fit."

y_hat <- fitted(model_1)
plot(x = y_hat, y = data$average_playtime, main = "Response vs. Fitted", xlab = "Fitted", ylab = "Average Playtime")
abline(a=0, b=1, lty=2)
```
To check the conditional mean response assumption, we analyze the Response vs. Fitted scatterplot @fig-resp-fit. The data points are tightly clustered in the bottom-left corner, which satisfies this assumption. The fitted line generally follows the line $y=hat{y}$, but there are notable outliers along the Y-axis where the observed `average_playtime` deviates significantly from the predicted values. These extreme outliers suggest that the response variable may be highly skewed or contain problematic data points.

To check the conditional mean predictor assumption, we examine the pairwise scatterplots of predictors @fig-pairwise. A lack of curves in all pairs of plots indicates that this assumption is satisfied. However, the Residuals vs. Fitted scatterplot @fig-resfit shows a fanning pattern resembling a left-pointing arrow, which suggests a potential violation of the constant variance assumption. Additionally, the Residuals vs. Qualitative Predictor scatterplots @fig-respre indicate possible violations of linearity and uncorrelated errors due to clusters of points and curves in the `positive_ratings` and `negative_ratings` predictors.

<!-- ## Create Pairwise Scatterplots for Predictors to look for multicollinearity -->
```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-pairwise
#| fig-cap: "Pairwise plot between numeric predictors"

ggpairs(data[, c(5, 6, 7, 11)]) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
        axis.text.y = element_text(size = 8))

```

<!-- ## Create Residuals vs Fitted Scatterplot to check Linearity, Uncorrelated Errors, and Constant Variance -->
```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-resfit
#| fig-cap: "Residual vs fitted plot for model lm(average_playtime ~ price + positive_ratings + 
#|  negative_ratings + achievements + owners, data=data)"

y_value <- resid(model_1)
x_value <- fitted(model_1)

plot(x = x_value, y = y_value, main = "Residual vs. Fitted", xlab = "Fitted", ylab = "Residuals")
abline(h = 0, col = "red", lty = 2)
```

Boxplots were used to assess linearity violations in the categorical predictor `owners`, and no issues were detected  @sec-model-detail. However, the Normal QQ Plot @fig-qq reveals significant deviations from the diagonal at the extremes, indicating a violation of the normality assumption.

The presence of these assumption violations aligns with the nature of the data, as confirmed by the outlier analysis in appendix @sec-outlier, which identifies several influential points even in the final model. For instance, the uncorrelated errors assumption may be violated because some predictors are inherently related (e.g., price and developer). These findings are consistent with those in the literature, where similar violations and errors have been reported. Addressing these violations through transformations is critical to ensure the stability of the model.

<!-- ## Create Normal QQ Plot to check Normality assumption -->
```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-qq
#| fig-cap: "Residual vs fitted plot for model"

qqnorm(y_value)
qqline(y_value)
```
<!-- ## Box-Cox Transformation -->
```{r}
#| echo: false
#| warning: false
#| message: false

data$price <- data$price + 1
data$positive_ratings <- data$positive_ratings + 1
data$negative_ratings <- data$negative_ratings + 1
data$achievements <- data$achievements + 1

# Define the Box-Cox transformation function
boxcox_transform <- function(x) {
  bc <- BoxCoxTrans(x)
  predict(bc, x)
}
# Apply the transformation to each predictor and the response
bc_average_playtime <- boxcox_transform(data$average_playtime)
bc_price <- boxcox_transform(data$price)
bc_positive_ratings <- boxcox_transform(data$positive_ratings)
bc_negative_ratings <- boxcox_transform(data$negative_ratings)
bc_achievements <- boxcox_transform(data$achievements)

```
<!-- ## Fitting transformed model -->
```{r}
#| echo: false
#| warning: false
#| message: false

# Load bc transformed model
model_transformed <- readRDS(file = here::here("models/model_bc_trans.rds"))
```

To mitigate these violations, Box-Cox transformations were applied to all numeric predictors to improve normality and address linearity issues. Different transformations were applied to the response and predictors. For instance, the response variable `average_playtime` was log-transformed after calculating a Box-Cox $\lambda$ 0.1, as an example, an original average_playtime value of 17,612 was transformed to 9.78. Note that a constant of 1 was added to all predictors before applying the Box-Cox transformation, as it requires strictly positive values.

After transforming the model, the Variance Inflation Factor (VIF) was calculated to check for multicollinearity. Notably, `bc_positive_ratings` (log-transformed `positive_ratings`) and `bc_negative_ratings` (log-transformed `negative_ratings`) exhibited moderately high VIF values of approximately 5 and 4.5 respectively, indicating multicollinearity. This is reasonable, as a higher number of positive reviews typically corresponds to greater attention and a larger player base of a game, which can also result in more negative reviews. The pairwise scatterplots @fig-pairwise confirm this, showing a positive linear correlation between these two predictors.

<!-- ## Checking for Multicollinearity -->
```{r}
#| include: false
#| echo: false
#| warning: false
#| message: false

vif(model_transformed)
```
To address this issue, `bc_negative_ratings` was removed from the model, leaving the final predictors as `bc_price` (log-transformed of `price`), `bc_positive_ratings`(log-transformed of `positive_ratings`),  `bc_achievements`(log-transformed of `achievements`) and `owners`. The refined model was re-evaluated with summary statistics and residual plots. The summary of coefficients as shown in @tbl-summary will be discussed in Section @sec-conclusion. The residual diagnostics @fig-reduced show significant improvement, with a more random distribution around zero in the Residuals vs. Fitted plot and a QQ plot with better alignment to the reference line.

<!-- ## Generatingn Residual Plots -->
```{r, fig.width=8, fig.height=6}
#| echo: false
#| warning: false
#| message: false
#| label: fig-reduced
#| fig-cap: "Diagnostic plots for lm(bc_average_playtime ~ bc_price + bc_positive_ratings + bc_achievements + owners, data=data)"

reduced <- readRDS(file = here::here("models/model_bc_reduced.rds"))
par(mfrow = c(2, 2))
plot(reduced)
# y_value <- resid(reduced)
# x_value <- fitted(reduced)

# plot(x = x_value, y = y_value, main = "Residual vs. Fitted", xlab = "Fitted", ylab = "Residuals")
```
To further simplify the model, automated selection procedures using AIC in both forward and backward directions were conducted. The procedure resulted in the same final model as described above, confirming its robustness.

<!-- ## AIC Model Selection -->
```{r}
#| include: false
#| echo: false
#| warning: false
#| message: false

step_model <- stepAIC(reduced, direction = "both")
summary(step_model)
```
<!-- ## ANOVA Comparison between full model and reduced model -->
```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false

anova(step_model, reduced)
# Conclude that step_model (without bc_achievements is better suited and no significant improvement to model from bc_achievements)
summary(step_model)$adj.r.squared
summary(reduced)$adj.r.squared
AIC(step_model)
AIC(reduced)
```

# Conclusion and Limitations {#sec-conclusion}

In conclusion, the model demonstrates that the average playtime of a game is positively influenced by several factors, including the game's price, the number of achievements, the number of positive reviews, and the estimated number of owners. Notably, the number of owners has the most substantial effect, as shown in the model summary @tbl-summary. For instance, games with an estimated ownership level of 100,000–200,000 have a coefficient of approximately 0.81; as ownership grows by a factor of 10, the effect increases to ~1.70, and for another 10-fold increase, the effect grows to ~3.04. This indicates that games with a larger player base tend to have higher average playtime, likely due to increased community engagement and multiplayer opportunities.

The positive relationship between price and average playtime is also significant. This is consistent with the notion that higher-priced games often provide more content, greater replayability, and better quality, which encourages players to spend more time playing. Additionally, players who invest more money in a game may feel a stronger obligation to "get their money's worth," resulting in longer playtime. Similarly, a greater number of positive reviews likely attracts more players and increases playtime, as positive feedback signals quality and encourages engagement.

Overall, the model aligns well with practical expectations. It suggests that developers and publishers aiming to increase average playtime should focus on improving game quality, fostering positive user reviews, and building a strong player base.

The limitations in this study is ...

# Ethics discussion
<!-- The Steam Games dataset [@davis2023steam] used in this research is a collected dataset, not simulated, comprising information on over 97,000 games published on the Steam platform. The metadata is comprehensively filled out, including details such as game titles, release dates, genres, and user ratings, which enhances the dataset's usability and reliability. The source of the data is clearly described, originating from the Steam store pages and Steam API, ensuring transparency and traceability. Additionally, the dataset is hosted on Kaggle, a reputable platform for data science and machine learning, and has been vetted and used in existing literature (discussed in the introduction). This widespread use and accessibility suggest that the dataset is well-regarded and trusted by third parties, further validating its credibility through the framework outlined in the ethics module. There are no ethical concerns to report. -->

\newpage

\appendix

# Appendix {-}

# Model details {#sec-model-detail}
## Residuals vs each Quantitative Predictor scatterplot to check Linearity, Uncorrelated Errors, and Constant Variance
```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-respre
#| fig-cap: "Residuals vs each Quantitative Predictor"

par(mfrow=c(1,4))

plot(x = data$price, y = y_value, main = "Residuals vs. Price", xlab = "Price", ylab = "Residuals")

plot(x = data$positive_ratings, y = y_value, main = "Residuals vs. Positive Ratings", xlab = "Positive Ratings", ylab = "Residuals")

plot(x = data$negative_ratings, y = y_value, main = "Residuals vs. Negative Ratings", xlab = "Negative Ratings", ylab = "Residuals")

plot(x = data$achievements, y = y_value, main = "Residuals vs. Achievements", xlab = "Achievements", ylab = "Residuals")
```

## BoxPlot to check assumptions for Categorical Predictors
See figure @fig-boc-cat
```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-boc-cat
#| fig-cap: "Boxplot residual vs. Owner levels suggest no linear assumption violation regarding the categorical predictors"

# par(mfrow=c(1,2))

# boxplot(y_value ~ data$developer , main="Residual vs Developer",
# xlab="Developer", ylab="Residuals")

boxplot(y_value ~ data$owners , main="Residual vs Owners",
xlab="Owners", ylab="Residuals")
```
## Final model results
<!-- ## Fitting Reduced Model -->
```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-summary

# Third model with reduce variable lm(bc_average_playtime ~ bc_price + bc_positive_ratings + bc_achievements + owners, data=data)
kable(tidy(reduced), caption = "Summary of the Reduced Model")
```

## Problematic points detection {#sec-outlier}

Checking for problematic points-leverage cutoff: ~0.006, cook's distance cutoff: ~0.961, dffits cutoff: ~0.105, dfbeta cutoff: ~0.025. Number of influential points by each case-leverage points: 216, regression outliers: 7, Cook's distance: 0, DFFITS: 211, by Beta from 0 to 7: 431, 476, 377, 394, 464, 321, 49, 0.
```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false

n <- nrow(data)
p <- length(coef(reduced)-1)
# leverage cutoff
hcut <- 2 * ((p + 1) / n)

# cook's distance cutoff
cookcut <- qf(0.5, df1 = p + 1, df2 = n - p - 1)

# dffits cutoff
fitcut <- 2 * sqrt((p + 1) / n)

# dfbeta cutoff
betacut <- 2 / sqrt(n)
# leverage statistic
h_ii <- hatvalues(reduced)

# standardized residuals
r_i <- rstandard(reduced)

# cook's distance 
D_i <- cooks.distance(reduced)

# dffits
dffits_i <- dffits(reduced)

# dfbetas
dfbetas_i <- dfbetas(reduced)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false

# Count the number of leverage points
leverage_points <- which(h_ii > hcut)

# Count the number of regression outliers
regression_outliers <- which(r_i > 4 | r_i < -4)

# Count the number of influential observations by Cook's distance
influential_cooks <- which(D_i > cookcut)

# Count the number of influential observations by DFFITS
influential_dffits <- which(abs(dffits_i) > fitcut)

# Count the number of influential observations by DFBETAS
for (i in 1:8) {
  beta_influential <- which(abs(dfbetas_i[, i]) > betacut)
  cat(paste0("  Beta ", i - 1, ": ", length(beta_influential), "\n"))
}
```

\newpage

# References