### 2024-12-04 paper/paper.Rmd
Rendered paper into pdf.

---
title: "Exploring Factors Influencing Average Playtime of Video Games: A Regression Analysis"
subtitle: ""
author: 
  - name: "Ziqi Zhu"
  - name: "Raghav Sinha"
  - name: "Christophe McWatt"
thanks: "Code and data are available at: [https://github.com/zzq20010617/sta302Paper](https://github.com/zzq20010617/sta302Paper)."
date: today
date-format: long
abstract: ""
number-sections: true
bibliography: references.bib
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| include: false
#| warning: false
#| message: false

library(ggplot2)
library(GGally)
library(readr)
library(tibble)
library(knitr)
library(kableExtra)
library(dplyr)
library(MASS)
library(caret)
library(car)
library(broom)
```

```{r}
#| include: false
#| warning: false
#| message: false

# Load analysis data
data <- read.csv(here::here("data/analysis_data/analysis_data.csv"))
```

# Contributions

We found the dataset together, and seperate the works as follow: Ziqi:
Introduction and data description Raghav: Ethics discussion and some
plot in preliminary result Christophe: Most of Preliminary results
section

# Introduction

The gaming industry has experienced exponential growth in recent years,
with millions of players engaging in various games on platforms like
Steam. Understanding what influences average playtime in games is
crucial for game developers, publishers, and marketers. Average playtime
is a key indicator of player engagement and game success, and
investigating factors that impact it can provide valuable insights into
game design, pricing strategies, and marketing efforts. So our research
aims to explore the factors that influence the average playtime of games
available on Steam using a dataset we get from Kaggle. Specifically, we
will investigate how variables such as price, user ratings, and other
game characteristics contribute to player engagement, as measured by
playtime. From some previous papers, we learned that both positive and
negative reviews have a significant correlation with playtime, but the
effect is more pronounced for positive reviews [@paper1]. Another paper
found out that game pricing significantly affects player engagement,
with lower-priced games generally seeing higher playtime [@paper2]. The
third paper did an empirical study that also examined the influence of
game reviews on playtime, focusing on how the number and sentiment of
reviews can affect player engagement. It concludes that games with more
positive reviews see increased playtime, especially if those reviews
highlight the game’s quality [@paper3]. We found this problem suitable
for linear regression because it allows us to model the relationship
between average playtime (the dependent variable) and various predictors
(price, ratings, developer, etc.). This approach helps quantify how each
factor influences playtime, enabling a clear understanding of the impact
of different game attributes. By fitting a linear regression model, we
can determine which factors have statistically significant effects on
playtime, providing actionable insights for game developers and
marketers.

# Methods

To develop and refine the final model for predicting average playtime, a
structured approach was employed that combined diagnostic assessments,
transformations, and model selection techniques. The analysis began with
data preparation and selection of relevant predictors based on domain
knowledge and preliminary exploration. A multiple linear regression
model was initially fitted using predictors including price, positive
ratings, negative ratings, number of achievements as numeric variables,
and ownership levels as categorical variable.

Preliminary model diagnostics, since we are using MLR model, we first
need to check conditional mean condition with both response and
predictors, including response vs. fitted plots and pairwise plots
between predictors. Then residuals vs. fitted plots were used to assess
the assumptions of linearity, constant-variance, and uncorrelated
errors. While Normal Quantile-Quantile (QQ) plot is used to check for
normality. To address violations of these assumptions, Box-Cox
transformations were applied to the response variable and select
predictors to make the most suited transform function based on their
$\lambda$. Then pairwise scatterplots and the Variance Inflation Factor
(VIF) were used to check for multicollinearity, and highly correlated
variables were removed or adjusted to improve the model's stability.

Refinement of the model was carried out through stepwise selection using
AIC (Akaike Information Criterion) to identify the best combination of
predictors while balancing model complexity and fit. ANOVA tests were
used to compare nested models and evaluate whether reductions in
predictors led to significant changes in explanatory power. The final
model was validated using diagnostic plots to confirm adherence to
linear regression assumptions, ensuring that it was suitable for
interpretation and prediction.

The methodology was implemented in programming language **R** [@citeR].
Packages **readr** [@readr], **httr** [@httr], **jsonlite** [@jsonlite]
were used to download the data, **dplyr** [@dplyr] and **lubridate**
[@lubridate] were used to clean the data. The diagnostic process is done
with help of **ggplot2** [@ggplot2] for residual graphs, **GGally**
[@GGally] and **car** [@car] for multicollinearity checks, **MASS**
[@MASS] for stepwise selection, **caret** [@caret] is used to perform
box-cox transform on variables. This systematic approach ensured that
each decision was guided by statistical evidence and diagnostic tools,
allowing for the development of a robust final model. A simplified
version of the methodology flowchart is presented in @fig-flowchart.

```{r}
#| warning: false
#| message: false
#| echo: false
#| out.width: 100%
#| label: fig-flowchart
#| fig-cap: "Methodology flowchart"

include_graphics(here::here("other/flowchart.png"))
```

<!-- ## Fit multiple linear regression model -->

```{r}
#| echo: false
#| warning: false
#| message: false

# Load first model fit with lm(average_playtime ~ price + positive_ratings + 
# negative_ratings + achievements + owners, data=analysis_data)
model_1 <- readRDS(file = here::here("models/model_1.rds"))
```

# Results

At the starting point, our baseline model uses `average_playtime` as the
response variable, with `price`, `positive_ratings`, `negative_ratings`,
`achievements`, and `owners` as predictors. The four key linear
regression assumptions considered are linearity, uncorrelated errors,
constant error variance, and normality. Additionally, since this is a
multiple linear regression model, we also examine the conditional mean
response and predictor assumptions.
<!-- ## Create Response vs Fitted Scatterplot to check patterns in residuals before further analysis. -->

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-resp-fit
#| fig-cap: "Response vs. Fitted plot to assess model fit."

y_hat <- fitted(model_1)
plot(x = y_hat, y = data$average_playtime, main = "Response vs. Fitted", xlab = "Fitted", ylab = "Average Playtime")
abline(a=0, b=1, lty=2)
```

To check the conditional mean response assumption, we analyze the
Response vs. Fitted scatterplot @fig-resp-fit. The data points are
tightly clustered in the bottom-left corner, which satisfies this
assumption. The fitted line generally follows the line $y=\hat{y}$, but
there are notable outliers along the Y-axis where the observed
`average_playtime` deviates significantly from the predicted values.
These extreme outliers suggest that the response variable may be highly
skewed or contain problematic data points.

To check the conditional mean predictor assumption, we examine the
pairwise scatterplots of predictors @fig-pairwise. A lack of curves in
all pairs of plots indicates that this assumption is satisfied. However,
the Residuals vs. Fitted scatterplot @fig-resfit shows a fanning pattern
resembling a left-pointing arrow, which suggests a potential violation
of the constant variance assumption. Additionally, the Residuals vs.
Qualitative Predictor scatterplots @fig-respre indicate possible
violations of linearity and uncorrelated errors due to clusters of
points and curves in the `positive_ratings` and `negative_ratings`
predictors.

<!-- ## Create Pairwise Scatterplots for Predictors to look for multicollinearity -->

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-pairwise
#| fig-cap: "Pairwise plot between numeric predictors"

ggpairs(data[, c(5, 6, 7, 11)]) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
        axis.text.y = element_text(size = 8))

```

<!-- ## Create Residuals vs Fitted Scatterplot to check Linearity, Uncorrelated Errors, and Constant Variance -->

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-resfit
#| fig-cap: "Residual vs fitted plot for model lm(average_playtime ~ price + positive_ratings + 
#|  negative_ratings + achievements + owners, data=data)"

y_value <- resid(model_1)
x_value <- fitted(model_1)

plot(x = x_value, y = y_value, main = "Residual vs. Fitted", xlab = "Fitted", ylab = "Residuals")
abline(h = 0, col = "red", lty = 2)
```

Boxplots were used to assess linearity violations in the categorical
predictor `owners`, and no issues were detected @sec-model-detail.
However, the Normal QQ Plot @fig-qq reveals significant deviations from
the diagonal at the extremes, indicating a violation of the normality
assumption.

The presence of these assumption violations aligns with the nature of
the data, as confirmed by the outlier analysis in appendix @sec-outlier,
which identifies several influential points even in the final model. For
instance, the uncorrelated errors assumption may be violated because
some predictors are inherently related (e.g., price and developer).
These findings are consistent with those in the literature, where
similar violations and errors have been reported. Addressing these
violations through transformations is critical to ensure the stability
of the model.

<!-- ## Create Normal QQ Plot to check Normality assumption -->

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-qq
#| fig-cap: "Residual vs fitted plot for model"

qqnorm(y_value)
qqline(y_value)
```

<!-- ## Box-Cox Transformation -->

```{r}
#| echo: false
#| warning: false
#| message: false

data$price <- data$price + 1
data$positive_ratings <- data$positive_ratings + 1
data$negative_ratings <- data$negative_ratings + 1
data$achievements <- data$achievements + 1

# Define the Box-Cox transformation function
boxcox_transform <- function(x) {
  bc <- BoxCoxTrans(x)
  predict(bc, x)
}
# Apply the transformation to each predictor and the response
bc_average_playtime <- boxcox_transform(data$average_playtime)
bc_price <- boxcox_transform(data$price)
bc_positive_ratings <- boxcox_transform(data$positive_ratings)
bc_negative_ratings <- boxcox_transform(data$negative_ratings)
bc_achievements <- boxcox_transform(data$achievements)

```

<!-- ## Fitting transformed model -->

```{r}
#| echo: false
#| warning: false
#| message: false

# Load bc transformed model
model_transformed <- readRDS(file = here::here("models/model_bc_trans.rds"))
```

To mitigate these violations, Box-Cox transformations were applied to
all numeric predictors to improve normality and address linearity
issues. Different transformations were applied to the response and
predictors. For instance, the response variable `average_playtime` was
log-transformed after calculating a Box-Cox $\lambda$ 0.1, as an
example, an original average_playtime value of 17,612 was transformed to
9.78. Note that a constant of 1 was added to all predictors before
applying the Box-Cox transformation, as it requires strictly positive
values.

After transforming the model, the Variance Inflation Factor (VIF) was
calculated to check for multicollinearity. Notably,
`bc_positive_ratings` (log-transformed `positive_ratings`) and
`bc_negative_ratings` (log-transformed `negative_ratings`) exhibited
moderately high VIF values of approximately 5 and 4.5 respectively,
indicating multicollinearity. This is reasonable, as a higher number of
positive reviews typically corresponds to greater attention and a larger
player base of a game, which can also result in more negative reviews.
The pairwise scatterplots @fig-pairwise confirm this, showing a positive
linear correlation between these two predictors.

<!-- ## Checking for Multicollinearity -->

```{r}
#| include: false
#| echo: false
#| warning: false
#| message: false

vif(model_transformed)
```

To address this issue, `bc_negative_ratings` was removed from the model,
leaving the final predictors as `bc_price` (log-transformed of `price`),
`bc_positive_ratings`(log-transformed of `positive_ratings`),
`bc_achievements`(log-transformed of `achievements`) and `owners`. The
refined model was re-evaluated with summary statistics and residual
plots. The summary of coefficients as shown in @tbl-summary will be
discussed in Section @sec-conclusion. The residual diagnostics
@fig-reduced show significant improvement, with a more random
distribution around zero in the Residuals vs. Fitted plot and a QQ plot
with better alignment to the reference line.

<!-- ## Generatingn Residual Plots -->

```{r, fig.width=8, fig.height=6}
#| echo: false
#| warning: false
#| message: false
#| label: fig-reduced
#| fig-cap: "Diagnostic plots for lm(bc_average_playtime ~ bc_price + bc_positive_ratings + bc_achievements + owners, data=data)"

reduced <- readRDS(file = here::here("models/model_bc_reduced.rds"))
par(mfrow = c(2, 2))
plot(reduced)
# y_value <- resid(reduced)
# x_value <- fitted(reduced)

# plot(x = x_value, y = y_value, main = "Residual vs. Fitted", xlab = "Fitted", ylab = "Residuals")
```

To further simplify the model, automated selection procedures using AIC
in both forward and backward directions were conducted. The procedure
resulted in the same final model as described above, confirming its
robustness.

<!-- ## AIC Model Selection -->

```{r}
#| include: false
#| echo: false
#| warning: false
#| message: false

step_model <- stepAIC(reduced, direction = "both")
summary(step_model)
```

<!-- ## ANOVA Comparison between full model and reduced model -->

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false

anova(step_model, reduced)
# Conclude that step_model (without bc_achievements is better suited and no significant improvement to model from bc_achievements)
summary(step_model)$adj.r.squared
summary(reduced)$adj.r.squared
AIC(step_model)
AIC(reduced)
```

# Conclusion and Limitations {#sec-conclusion}

In conclusion, the model demonstrates that the average playtime of a
game is positively influenced by several factors, including the game's
price, the number of achievements, the number of positive reviews, and
the estimated number of owners. Notably, the number of owners has the
most substantial effect, as shown in the model summary @tbl-summary. For
instance, games with an estimated ownership level of 100,000–200,000
have a coefficient of approximately 0.81; as ownership grows by a factor
of 10, the effect increases to \~1.70, and for another 10-fold increase,
the effect grows to \~3.04. This indicates that games with a larger
player base tend to have higher average playtime, likely due to
increased community engagement and multiplayer opportunities.

The positive relationship between price and average playtime is also
significant. This is consistent with the notion that higher-priced games
often provide more content, greater replayability, and better quality,
which encourages players to spend more time playing. Additionally,
players who invest more money in a game may feel a stronger obligation
to "get their money's worth," resulting in longer playtime. Similarly, a
greater number of positive reviews likely attracts more players and
increases playtime, as positive feedback signals quality and encourages
engagement.

Overall, the model aligns well with practical expectations. It suggests
that developers and publishers aiming to increase average playtime
should focus on improving game quality, fostering positive user reviews,
and building a strong player base.

This study faces limitations primarily due to assumption violations and
data issues. The Residuals vs. Fitted plot revealed a fanning pattern,
indicating heteroscedasticity, while the Normal QQ Plot showed
significant non-normality. Outliers in the response variable
`average_playtime` suggest skewed data, despite transformations like
Box-Cox improving stability. Multicollinearity between
`bc_positive_ratings` and `bc_negative_ratings` required removing the
latter, though some interdependence remains.

The dataset, limited to Steam games, introduces platform-specific
biases, restricting generalizability. Observational data prevents causal
inference, and omitted variables, such as marketing strategies, likely
impact playtime. Future studies should refine transformations, address
multicollinearity more rigorously, and expand variables and platforms to
enhance the robustness and applicability of the findings.

# Ethics discussion

The Steam Games dataset [@davis2023steam] used in this research contains
metadata on over 97,000 games published on the Steam platform. The
dataset is sourced transparently from Steam store pages and the Steam
API, hosted on Kaggle, and widely used in existing literature, ensuring
its credibility.

In our analysis, we chose automated selection methods due to the
dataset's size and complexity. Manual selection methods were not used as
they would have been impractical and prone to human error, reducing
consistency and reproducibility. While both methods are ethically
equivalent when executed carefully, automated methods better align with
the ethical principles outlined in the second ethics module by
minimizing foreseeable risks and ensuring reliability.

Avoiding blameworthy practices requires addressing foreseeability,
avoidability, and potential errors. Automated methods allowed us to
mitigate these risks efficiently, ensuring accountability in our
approach. Manual methods, while suitable for smaller datasets, would
have introduced avoidable errors and diminished reliability, which could
be considered ethically negligent.

By prioritizing virtues such as precision and accountability, we ensured
an ethically sound methodology. Our choice was guided by practical
considerations without compromising ethical standards, fostering robust
and reproducible results.

\newpage

\appendix

# Appendix {.unnumbered}

# Model details {#sec-model-detail}

## Residuals vs each Quantitative Predictor scatterplot to check Linearity, Uncorrelated Errors, and Constant Variance

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-respre
#| fig-cap: "Residuals vs each Quantitative Predictor"

par(mfrow=c(1,4))

plot(x = data$price, y = y_value, main = "Residuals vs. Price", xlab = "Price", ylab = "Residuals")

plot(x = data$positive_ratings, y = y_value, main = "Residuals vs. Positive Ratings", xlab = "Positive Ratings", ylab = "Residuals")

plot(x = data$negative_ratings, y = y_value, main = "Residuals vs. Negative Ratings", xlab = "Negative Ratings", ylab = "Residuals")

plot(x = data$achievements, y = y_value, main = "Residuals vs. Achievements", xlab = "Achievements", ylab = "Residuals")
```

## BoxPlot to check assumptions for Categorical Predictors

See figure @fig-boc-cat

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-boc-cat
#| fig-cap: "Boxplot residual vs. Owner levels suggest no linear assumption violation regarding the categorical predictors"

# par(mfrow=c(1,2))

# boxplot(y_value ~ data$developer , main="Residual vs Developer",
# xlab="Developer", ylab="Residuals")

boxplot(y_value ~ data$owners , main="Residual vs Owners",
xlab="Owners", ylab="Residuals")
```

## Final model results

<!-- ## Fitting Reduced Model -->

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-summary

# Third model with reduce variable lm(bc_average_playtime ~ bc_price + bc_positive_ratings + bc_achievements + owners, data=data)
kable(tidy(reduced), caption = "Summary of the Reduced Model")
```

## Problematic points detection {#sec-outlier}

Checking for problematic points-leverage cutoff: \~0.006, cook's
distance cutoff: \~0.961, dffits cutoff: \~0.105, dfbeta cutoff:
\~0.025. Number of influential points by each case-leverage points: 216,
regression outliers: 7, Cook's distance: 0, DFFITS: 211, by Beta from 0
to 7: 431, 476, 377, 394, 464, 321, 49, 0.

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false

n <- nrow(data)
p <- length(coef(reduced)-1)
# leverage cutoff
hcut <- 2 * ((p + 1) / n)

# cook's distance cutoff
cookcut <- qf(0.5, df1 = p + 1, df2 = n - p - 1)

# dffits cutoff
fitcut <- 2 * sqrt((p + 1) / n)

# dfbeta cutoff
betacut <- 2 / sqrt(n)
# leverage statistic
h_ii <- hatvalues(reduced)

# standardized residuals
r_i <- rstandard(reduced)

# cook's distance 
D_i <- cooks.distance(reduced)

# dffits
dffits_i <- dffits(reduced)

# dfbetas
dfbetas_i <- dfbetas(reduced)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false

# Count the number of leverage points
leverage_points <- which(h_ii > hcut)

# Count the number of regression outliers
regression_outliers <- which(r_i > 4 | r_i < -4)

# Count the number of influential observations by Cook's distance
influential_cooks <- which(D_i > cookcut)

# Count the number of influential observations by DFFITS
influential_dffits <- which(abs(dffits_i) > fitcut)

# Count the number of influential observations by DFBETAS
for (i in 1:8) {
  beta_influential <- which(abs(dfbetas_i[, i]) > betacut)
  cat(paste0("  Beta ", i - 1, ": ", length(beta_influential), "\n"))
}
```

\newpage

# References